Phenotypic evolution is less constrained in complex food webs or Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-09-20 ms_complexity_selection.Rmd
many many changes

---
title: Food-web complexity alters the fitness landscape of an insect herbivore
author:
  - name: Matthew A. Barbour
    email: matthew.barbour@ieu.uzh.ch
    affiliation: a,b
    footnote: Corresponding Author
  - name: Christopher J. Greyson-Gaito
    affiliation: a,c
  - name: Arezoo Sootodeh
    affiliation: a
  - name: Brendan Locke
    affiliation: d
  - name: Jordi Bascompte
    affiliation: b
address:
  - code: a
    address: University of British Columbia, Department of Zoology, 6270 University Blvd., Vancouver, BC, V6T 1Z4, Canada
  - code: b
    address: University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, Zurich, 8057, Switzerland
  - code: c
    address: University of Guelph, Department of Integrative Biology, 50 Stone Rd. East, Guelph, ONT, N1G 2W1, Canada
  - code: d
    address: Humboldt State University, Department of Biological Sciences, 1 Harpst St., Arcata, CA, 95521, USA
abstract: "Studies of natural selection and fitness landscapes usually treat the network of interacting species as a “black box”. Given that     the loss of biodiversity is simplifying the structure of ecological networks, there is a pressing need to answer the question: how does network complexity affect natural selection and the fitness landscape of associated species? To answer this question, we conducted a field experiment that manipulated the complexity of a food web associated with a galling insect herbivore. To maintain complex food webs, we allowed the entire community of natural enemies to attack insect galls on 64 plants in a common garden setting. To create simple food webs, we excluded a guild of three larval parasitoids by bagging galls on 64 different plants; therefore, mortality in this treatment was primarily due to a single egg parasitoid that attacks prior to gall formation. We then measured herbivore survival as a function of three key gall traits in each treatment. We found that more traits were under selection in the simple vs. complex food web. This occurred because different parasitoid species impose different selection pressures on gall traits, thereby minimizing relative fitness differences among insect galls with different phenotypes. Our work suggests that more complex food webs allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution to environmental change."

@@

<!-- Species-interaction networks provide a mechanistic link between community and population ecology, as they describe who interacts with whom in a community. Similarly, fitness landscapes provide a mechanistic link between population and evolutionary ecology, as they describe the influence of natural selection on phenotypic variation. It remains unclear, however, how the network structure of species interactions shapes the fitness landscape. Understanding this relationship is likely key for developing a predictive ecology across scales of biological organization. To examine the relationship between network structure and the fitness landscape, we conducted a field experiment that manipulated the network of trophic interactions (simple vs. complex) influencing the fitness of an insect herbivore. We then quantified the fitness landscape of herbivores in each treatment by measuring herbivore survival as a function of multiple phenotypic traits. We found that more traits were under selection in the simple vs. complex trophic network. This occurred because different natural enemies impose different selection pressures on herbivore traits, thereby minimizing relative fitness differences among herbivore phenotypes in the complex network. Our work suggests that more complex trophic networks allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution of populations to environmental change.-->

@@

## Load required libraries ----
library(tidyverse)
library(cowplot) # pretty default ggplots
#library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
#library(gamm4) # generalized additive mixed-models
library(gsg) # calculating selection gradients
#library(visreg)
library(viridis) # for color palette
library(lme4) # for generalized linear mixed models

@@

  # subset data for analysis
  filter(phenology == "early", Location == "tree",
         platy > 0 | ectos > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  mutate(gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg_parasitoid = as.numeric(ifelse(platy > 0, 1, 0)),
         sc.gall_size = as.numeric(scale(Gall_Height_mm)), 
         sc.clutch_size = as.numeric(scale(gall_individuals)), 
         sc.log.clutch_size = as.numeric(scale(log(gall_individuals))), 
         sc.female_preference = as.numeric(scale(Density_per_100_shoots)),
         sc.log1p.female_preference = as.numeric(scale(log1p(Density_per_100_shoots))))

## Create separate datasets for quantifying selection gradients ----
control_df <- gall_selection.df %>%
  filter(Treatment.focus == "Control") #%>%
  #mutate(sc.gall_size = as.numeric(scale(Gall_Height_mm)), 
  #       sc.clutch_size = as.numeric(scale(gall_individuals)), 
  #       sc.log.clutch_size = as.numeric(scale(log(gall_individuals))), 
  #       sc.female_preference = as.numeric(scale(Density_per_100_shoots)),
  #       sc.log1p.female_preference = as.numeric(scale(log1p(Density_per_100_shoots))))

treatment_df <- gall_selection.df %>%
  filter(Treatment.focus == "Ectoparasitoid exclusion", ectos < 1) #%>%  # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  #mutate(sc.gall_size = as.numeric(scale(Gall_Height_mm)), 
  #       sc.clutch_size = as.numeric(scale(gall_individuals)), 
  #       sc.log.clutch_size = as.numeric(scale(log(gall_individuals))), 
  #       sc.female_preference = as.numeric(scale(Density_per_100_shoots)),
  #       sc.log1p.female_preference = as.numeric(scale(log1p(Density_per_100_shoots))))

mean(filter(gall_selection.df, Treatment.focus == "Ectoparasitoid exclusion")$ectos) # less than 3% of larva were parasitized by an ectoparasitoid in the exclusion experiment.

@@

@@

## Calculate selection gradients following method of Janzen and Stern 1998, Evolution
Beta_avg_grad <- function(.) {
  mean_fitness <- mean(predict(., type = "response")) #inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness 
}

@@

@@

## Number of bootstrap replicates
n_boots_analysis <- 1000
n_boots_plots <- 100

@@

AIC comparison suggests gam is better than quadratic approximation; however, I'm worried that the GAM model is overfitting. I feel like the quadratic approximation, while not providing the best fit, will still preven the model from overfitting.

Now the question is: normal glm vs. mixed-model? Maybe I need to answer, why not the mixed model?

- There is no theory developed for them. 
- Relative fitness occurs at different hierarchical levels instead of the level of the sampled population.
- Less straightforward.

Mixed-models are also good for modeling heterogeneity, but if I'm interested in a general population level, then maybe I don't have to worry about that. 
```{r GAM, include=F}
control_gam <- gam(gall_survival ~ s(sc.gall_size) + s(sc.log.clutch_size, k=9) + s(sc.log1p.female_preference), control_df, family=binomial(link=logit))
plot(control_gam, trans = inverse_logit)

treatment_gam <- gam(gall_survival ~ s(sc.gall_size) + s(sc.log.clutch_size, k=9) + s(sc.log1p.female_preference), treatment_df, family=binomial(link=logit))
plot(treatment_gam, trans = inverse_logit)

control_gam_quad <- gam(gall_survival ~ sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2), control_df, family=binomial(link=logit))

treatment_gam_quad <- gam(gall_survival ~ sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2), treatment_df, family=binomial(link=logit))

AIC(control_gam, control_gam_quad)
AIC(treatment_gam, treatment_gam_quad)

hist(gall_selection.df$sc.gall_size)
hist(gall_selection.df$sc.log.clutch_size)
hist(gall_selection.df$sc.log1p.female_preference)

control_glm <- glm(gall_survival ~ sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2) + sc.gall_size:sc.log.clutch_size + sc.gall_size:sc.log1p.female_preference + sc.log.clutch_size:sc.log1p.female_preference, control_df, family=quasibinomial())#family=binomial(link=logit))
summary(control_glm)
visreg::visreg(control_glm, scale = "response")

treatment_glm <- glm(gall_survival ~ sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2) + sc.gall_size:sc.log.clutch_size + sc.gall_size:sc.log1p.female_preference + sc.log.clutch_size:sc.log1p.female_preference, treatment_df, family=quasibinomial())#family=binomial(link=logit))
summary(treatment_glm)

visreg::visreg(control_glm, scale = "response")
visreg::visreg(treatment_glm, scale = "response")

test_glm <- glm(gall_survival ~ Treatment.focus*(sc.gall_size + I(sc.gall_size^2) + sc.log.clutch_size + I(sc.log.clutch_size^2) + sc.log1p.female_preference + I(sc.log1p.female_preference^2) + sc.gall_size:sc.log.clutch_size + sc.gall_size:sc.log1p.female_preference + sc.log.clutch_size:sc.log1p.female_preference), gall_selection.df, family=quasibinomial())#family=binomial(link=logit))
summary(test_glm)

visreg::visreg2d(test_glm, xvar="sc.gall_size", yvar="sc.log.clutch_size", scale="response")
```

```{r, Cryptic effects of parasitism on gall morphology, include=F}
# how does this analysis compare to Heath?
# does it make sense to look for the interaction?

control_gall_size <- lmer(sc.gall_size ~ gall_survival +
                           (1|Genotype:Plant_Position:Gall_Number),
                         data = control_df)
summary(control_gall_size)
visreg::visreg(control_gall_size, xvar="gall_survival")

piecewiseSEM::rsquared(control_gall_size)

flip.control_gall_size <- lmer(gall_survival ~ sc.gall_size +
                           (1|Genotype:Plant_Position:Gall_Number),
                         data = control_df)
summary(flip.control_gall_size)
visreg::visreg(flip.control_gall_size, xvar="sc.gall_size")

lm.control_gall_size <- lm(Gall_Height_mm ~ gall_survival,
                         data = control_df)
summary(lm.control_gall_size)
visreg::visreg(lm.control_gall_size, xvar="gall_survival")

lmer.control_gall_size <- lmer(Gall_Height_mm ~ gall_survival + (1|Genotype:Plant_Position:Gall_Number),
                         data = control_df)
summary(lmer.control_gall_size)
visreg::visreg(lmer.control_gall_size, xvar="gall_survival")

lmer.control_gall_size_2 <- lmer(Gall_Height_mm ~ gall_survival + (1|Genotype/Plant_Position/Gall_Number),
                         data = control_df)
summary(lmer.control_gall_size_2)
visreg::visreg(lmer.control_gall_size_2, xvar="gall_survival")

piecewiseSEM::rsquared(flip.control_gall_size)

treatment_gall_size <- lmer(sc.gall_size ~ gall_survival +
                           (1|Genotype:Plant_Position:Gall_Number),
                         data = treatment_df)
summary(treatment_gall_size)
visreg::visreg(treatment_gall_size, xvar="gall_survival")

lm.treatment_gall_size <- lm(Gall_Height_mm ~ gall_survival,
                         data = treatment_df)
summary(lm.treatment_gall_size)
visreg::visreg(lm.treatment_gall_size, xvar="gall_survival")

lmer.treatment_gall_size <- lmer(Gall_Height_mm ~ gall_survival + (1|Genotype:Plant_Position:Gall_Number),
                         data = treatment_df)
summary(lmer.treatment_gall_size)
visreg::visreg(lmer.treatment_gall_size, xvar="gall_survival")

lmer.treatment_gall_size_2 <- lmer(Gall_Height_mm ~ gall_survival + (1|Genotype/Plant_Position/Gall_Number),
                         data = treatment_df)
summary(lmer.treatment_gall_size_2)
visreg::visreg(lmer.treatment_gall_size_2, xvar="gall_survival")

piecewiseSEM::rsquared(lmer.treatment_gall_size_2)

lmer.treatment_sc.gall_size_2 <- lmer(sc.gall_size ~ gall_survival + sc.clutch_size + sc.female_preference + (1|Genotype/Plant_Position/Gall_Number),
                         data = treatment_df)
summary(lmer.treatment_sc.gall_size_2)

lm.treatment_sc.gall_size <- lm(sc.gall_size ~ gall_survival,
                         data = treatment_df)
summary(lm.treatment_sc.gall_size)

summary(lm(gall_survival/mean(gall_survival) ~ sc.gall_size + sc.log.clutch_size + sc.log1p.female_preference + I(sc.gall_size^2) + I(sc.log.clutch_size^2) + I(sc.log1p.female_preference^2), data=treatment_df))
summary(lmer(gall_survival/mean(gall_survival) ~ sc.gall_size + sc.log.clutch_size + sc.log1p.female_preference + I(sc.gall_size^2) + I(sc.log.clutch_size^2) + I(sc.log1p.female_preference^2) + (1|Genotype/Plant_Position/Gall_Number), data=treatment_df))

summary(lm(gall_survival/mean(gall_survival) ~ sc.gall_size + sc.log.clutch_size + sc.log1p.female_preference + I(sc.gall_size^2) + I(sc.log.clutch_size^2) + I(sc.log1p.female_preference^2) , data=control_df))
summary(lmer(gall_survival/mean(gall_survival) ~ sc.gall_size + sc.log.clutch_size + sc.log1p.female_preference + I(sc.gall_size^2) + I(sc.log.clutch_size^2) + I(sc.log1p.female_preference^2) + (1|Genotype/Plant_Position/Gall_Number), data=control_df))

pair_control <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size)) %>%
  spread(gall_survival, mean_gall_size) %>%
  mutate(diff = `1` - `0`)
summary(pair_control$diff)
length(pair_control$diff)

t.test(pair_control$`1`, pair_control$`0`, paired = T)

pair_treatment <- treatment_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size)) %>%
  spread(gall_survival, mean_gall_size) %>%
  mutate(diff = `1` - `0`)
summary(pair_treatment$diff)
length(pair_treatment$diff)
pair_treatment

t.test(pair_treatment$`1`, pair_treatment$`0`, paired = T)

pair_control_ectos <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(ectos)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size)) %>%
  spread(gall_survival, mean_gall_size) %>%
  mutate(diff = `1` - `0`)
summary(pair_control_ectos$diff)
length(pair_control_ectos$diff)

t.test(pair_control_ectos$`1`, pair_control_ectos$`0`, paired = T)

pair_control_platy <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(egg_parasitoid)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size)) %>%
  spread(gall_survival, mean_gall_size) %>%
  mutate(diff = `1` - `0`)
summary(pair_control_platy$diff)
length(pair_control_platy$diff)

t.test(pair_control_platy$`1`, pair_control_platy$`0`, paired = T)

gsg::moments.differentials(z=control_df$sc.gall_size, W = control_df$gall_survival)
gsg::moments.differentials(z=treatment_df$sc.gall_size, W = treatment_df$gall_survival)

summary(lm(gall_survival/mean(gall_survival) ~ sc.gall_size, control_df))
summary(lm(gall_survival/mean(gall_survival) ~ sc.gall_size, treatment_df))

test_control_df <- bind_rows(mutate(control_df, type = 0), mutate(filter(control_df, gall_survival == 1), type = 1))
summary(lm(sc.gall_size ~ type, test_control_df)) # virtually the same as above

test_treatment_df <- bind_rows(mutate(treatment_df, type = 0), mutate(filter(treatment_df, gall_survival == 1), type = 1))
summary(lm(sc.gall_size ~ type, test_treatment_df)) # virtually the same as above

# can also look at it this way and just look at the mean difference before and after selection
test_control_df %>%
  group_by(type) %>%
  summarise(mean_gall_size = mean(sc.gall_size))

test_treatment_df %>%
  group_by(type) %>%
  summarise(mean_gall_size = mean(sc.gall_size))

# does this enable me to estimate the confounding effect of parasitism on gall size? 
# Note that the intercept is no longer centered on zero...
summary(lmer(sc.gall_size ~ type + (1|Genotype/Plant_Position/Gall_Number), test_control_df))
summary(lmer(sc.gall_size ~ type + (1|Genotype/Plant_Position/Gall_Number), test_treatment_df))

moments_control_df <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size))

gsg::moments.differentials(z=moments_control_df$mean_gall_size, W=moments_control_df$gall_survival) # does this mean anything?

moments_treatment_df <- treatment_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  group_by(Gall_Number, gall_survival) %>%
  summarise(mean_gall_size = mean(sc.gall_size))

gsg::moments.differentials(z=moments_treatment_df$mean_gall_size, W=moments_treatment_df$gall_survival)

### MOST PROMISING ANALYSIS THUS FAR THAT MAY ENABLE ME TO ASSESS THE BIAS OF PARASITISM ON OBSERVED SELECTION GRADIENTS.
test_again_control <- control_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  mutate(all_mean_gall_size = mean(sc.gall_size)) %>%
  filter(gall_survival == 1) %>%
  mutate(survive_mean_gall_size = mean(sc.gall_size)) %>%
  distinct(Gall_Number, all_mean_gall_size, survive_mean_gall_size) %>%
  mutate(diff_gall_size = survive_mean_gall_size - all_mean_gall_size)  # all of these represent selection differentials at the gall level

hist(test_again_control$diff_gall_size)
t.test(test_again_control$survive_mean_gall_size, test_again_control$all_mean_gall_size, paired = T)

test_ectos_control <- control_df %>%
  filter(pupa > 0 | ectos > 0) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  mutate(all_mean_gall_size = mean(sc.gall_size)) %>%
  filter(gall_survival == 1) %>%
  mutate(survive_mean_gall_size = mean(sc.gall_size)) %>%
  distinct(Gall_Number, all_mean_gall_size, survive_mean_gall_size) %>%
  mutate(diff_gall_size = survive_mean_gall_size - all_mean_gall_size)  # all of these represent selection differentials at the gall level

hist(test_ectos_control$diff_gall_size)
t.test(test_ectos_control$survive_mean_gall_size, test_ectos_control$all_mean_gall_size, paired = T)

test_platy_control <- control_df %>%
  filter(pupa > 0 | egg_parasitoid > 0) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  mutate(all_mean_gall_size = mean(sc.gall_size)) %>%
  filter(gall_survival == 1) %>%
  mutate(survive_mean_gall_size = mean(sc.gall_size)) %>%
  distinct(Gall_Number, all_mean_gall_size, survive_mean_gall_size) %>%
  mutate(diff_gall_size = survive_mean_gall_size - all_mean_gall_size)  # all of these represent selection differentials at the gall level

hist(test_platy_control$diff_gall_size)
t.test(test_platy_control$survive_mean_gall_size, test_platy_control$all_mean_gall_size, paired = T)

test_again_treatment <- treatment_df %>%
  #filter(gall_individuals==2) %>%
  group_by(Gall_Number) %>%
  mutate(mean_survival = mean(gall_survival)) %>%
  #filter(mean_survival == 0.5) %>%
  filter(mean_survival > 0, mean_survival < 1) %>%
  select(Gall_Number, mean_survival, gall_survival, sc.gall_size) %>%
  mutate(all_mean_gall_size = mean(sc.gall_size)) %>%
  filter(gall_survival == 1) %>%
  mutate(survive_mean_gall_size = mean(sc.gall_size)) %>%
  distinct(Gall_Number, all_mean_gall_size, survive_mean_gall_size) %>%
  mutate(diff_gall_size = survive_mean_gall_size - all_mean_gall_size)  # all of these represent selection differentials at the gall level

hist(test_again_treatment$diff_gall_size)
t.test(test_again_treatment$survive_mean_gall_size, test_again_treatment$all_mean_gall_size, paired = T)
```

```{r Calculate Mean and SD of Gall Traits at Different Scales, include=F, eval=F}

## Calculate mean and SD of gall traits at different scales ----
summary.individual_level <- gall_selection.df %>%
  group_by(Treatment.focus) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean, sd), na.rm = TRUE) %>%
  mutate(Level = "Individual")

mean.gall_size.control <- filter(summary.individual_level, Treatment.focus == "Control")$Gall_Height_mm_mean
sd.gall_size.control <- filter(summary.individual_level, Treatment.focus == "Control")$Gall_Height_mm_sd
mean.gall_size.treatment <- filter(summary.individual_level, Treatment.focus == "Ectoparasitoid exclusion")$Gall_Height_mm_mean
sd.gall_size.treatment <- filter(summary.individual_level, Treatment.focus == "Ectoparasitoid exclusion")$Gall_Height_mm_sd

summary.gall_level <- gall_selection.df %>%
  group_by(Treatment.focus, Gall_Number) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean), na.rm = TRUE) %>%
  group_by(Treatment.focus) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean, sd), na.rm = TRUE) %>%
  mutate(Level = "Gall")

mean.clutch_size.control <- filter(summary.gall_level, Treatment.focus == "Control")$gall_individuals_mean
sd.clutch_size.control <- filter(summary.gall_level, Treatment.focus == "Control")$gall_individuals_sd
mean.clutch_size.treatment <- filter(summary.gall_level, Treatment.focus == "Ectoparasitoid exclusion")$gall_individuals_mean
sd.clutch_size.treatment <- filter(summary.gall_level, Treatment.focus == "Ectoparasitoid exclusion")$gall_individuals_sd

summary.plant_level <- gall_selection.df %>%
  group_by(Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean), na.rm = TRUE) %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean), na.rm = TRUE) %>%
  group_by(Treatment.focus) %>%
  summarise_at(vars(Gall_Height_mm, gall_individuals, Density_per_100_shoots), funs(mean, sd), na.rm = TRUE) %>%
  mutate(Level = "Plant")

mean.female_preference.control <- filter(summary.plant_level, Treatment.focus == "Control")$Density_per_100_shoots_mean
sd.female_preference.control <- filter(summary.plant_level, Treatment.focus == "Control")$Density_per_100_shoots_sd
mean.female_preference.treatment <- filter(summary.plant_level, Treatment.focus == "Ectoparasitoid exclusion")$Density_per_100_shoots_mean
sd.female_preference.treatment <- filter(summary.plant_level, Treatment.focus == "Ectoparasitoid exclusion")$Density_per_100_shoots_sd

summary.gall_traits <- bind_rows(summary.individual_level, summary.gall_level, summary.plant_level)
```

<!-- > *There is a much more insidious kind of extinction: the extinction of ecological interactions.* 
> *(Janzen, D.H. 1974. The deflowering of Central America. Natural History. 83:48–53).* -->

@@

<!-- I. Fitness Landscape as a unifying framework for evolution and ecology at the population-level -->

@@

<!--Prior work with this study system has shown that there is directional selection for larger galls, likely because larger galls provide a refuge from parasitoid attack (@Barbour2016). 
However, there is also evidence that each parasitoid species imposes differential selection on gall traits (@Barbour2016).-->

@@

<!-- I. Evolutionary dynamics in species-rich communities remains unclear, especially in the field -->

<!-- Biological diversity -- from genes, to phenotypes, to species -- has and continues to be shaped by the interplay between ecological and evolutionary processes. <!-- need to improve this instead of just saying its fascinated a small group of people --> 
<!-- Much of this biological diversity has been molded by natural selection arising from species interactions, such as resource competition (@Schluter2000), mutualisms (@Jordano1987), and predation (@Abrams2000). 
While there is clear evidence that pairwise interactions can drive evolution, we also know that most species interact with multiple species in an ecological community. <!-- are embedded within species-rich communities.--> <!-- complex networks of interacting species. <!-- need to make the link to multiple traits and the adaptive landscape here --> <!-- Somthing about the this work remaining eminently theoretical --> 
<!-- Understanding how evolutionary dynamics unfold in a community context<!-- when moving beyond pairwise interactions--><!-- in more species-rich communities--><!--  is challenging and eminently theoretical (@McPeek2017, @De_Mazancourt2008, @Guimaraes2017, @Nuismer2013). 
<!-- Do I need to give more credit here to work by @TerHorst2018 and others?). I think it would be useful to show that there is some field and lab experiments demonstrating that community context can alter evolutionary trajectories -->
<!-- Given the rapid loss of species diversity we are experiencing throughout the world (cite), we are in urgent need of work that makes and tests predictions for how the loss of species will affect the evolutionary process in natural communities.

<!-- Theoretical models have begun to examine how the network structure of species interactions drives evolutionary change (Nuismer paper; Guimeras paper; Ecology Letters paper from a spanish guy...; @McPeek2017); however, we are currently lacking experimental tests of how changes in network structure will alter the dynamics of natural selection in the field. -->

<!-- II. Linking species loss to changes in network complexity -->

Predicting the evolutionary consequences of species loss first requires an understanding of the concommitant change in the species-interaction networks.<!-- the relationship between the network structure of a community and the fitness landscape of its constituent species.--> 
Knowing the interaction network is crucial because the loss of biodiversity, in and of itself, will not alter evolution -- it is the associated loss of ecological interactions that will affect evolutionary change (@Janzen1974). 
For a network of directly connected species, we would expect that the loss of species to result in a loss of network complexity.<!-- Need to formalize this logic more, because the loss of a specialist species in a highly generalist food web, would actually increase network complexity... Maybe I need to focus on the node/species level property --> 
Network complexity is a property that describes the diversity of interactions in an ecological community (@Banasek-Richter2009). 
Thus, all else equal, species loss will decrease the diversity of interactions, resulting in a more simple network. <!-- Explicitly linking network complexity to the fitness landscape, however, is ultimately needed to predict the evolutionary consequences of biodiversity loss. -->

>>>>

---
title: Phenotypic evolution is less constrained in complex food webs or Food-web complexity alters the fitness landscape of an insect herbivore 
author:
  - name: Matthew A. Barbour
    email: matthew.barbour@ieu.uzh.ch
    affiliation: a,b
    footnote: Corresponding Author
  - name: Christopher J. Greyson-Gaito
    affiliation: a,c
  - name: Arezoo Sootodeh
    affiliation: a
  - name: Brendan Locke
    affiliation: d
  - name: Jordi Bascompte
    affiliation: b
address:
  - code: a
    address: University of British Columbia, Department of Zoology, 6270 University Blvd., Vancouver, BC, V6T 1Z4, Canada
  - code: b
    address: University of Zurich, Department of Evolutionary Biology and Environmental Studies, Winterthurerstrasse 190, Zurich, 8057, Switzerland
  - code: c
    address: University of Guelph, Department of Integrative Biology, 50 Stone Rd. East, Guelph, ONT, N1G 2W1, Canada
  - code: d
    address: Humboldt State University, Department of Biological Sciences, 1 Harpst St., Arcata, CA, 95521, USA
abstract: "Species-interaction networks provide a mechanistic link between community and population ecology, as they describe who interacts with whom in a community. Similarly, fitness landscapes provide a mechanistic link between population and evolutionary ecology, as they describe the influence of natural selection on phenotypic variation. It remains unclear, however, how the network structure of species interactions shapes the fitness landscape. Understanding this relationship is likely key for developing a predictive ecology across scales of biological organization. To examine the relationship between network structure and the fitness landscape, we conducted a field experiment that manipulated the network of trophic interactions (simple vs. complex) influencing the fitness of an insect herbivore. We then quantified the fitness landscape of herbivores in each treatment by measuring herbivore survival as a function of multiple phenotypic traits. We found that more traits were under selection in the simple vs. complex trophic network. This occurred because different natural enemies impose different selection pressures on herbivore traits, thereby minimizing relative fitness differences among herbivore phenotypes in the complex network. Our work suggests that more complex trophic networks allow phenotypic variation to persist, which could facilitate subsequent adaptive evolution of populations to environmental change."

@@

@@

## Load required libraries ----
library(car)        # for homogeneity of variance test
library(tidyverse)  # for managing data
library(cowplot)    # pretty default ggplots
library(broom)      # for tidying multiple linear models
library(visreg)     # for quick visualizations of model effects
library(viridis)    # for color palette
library(lme4)       # for generalized linear mixed models
library(piecewiseSEM) # for calculating model R-square
library(latex2exp)  # using Latex notation for axes labels

@@

  # subset data for analysis
  filter(phenology == "early", Location == "tree",
         platy > 0 | ectos > 0 | pupa > 0) %>%                  # eliminate unknown sources of mortality
  filter(Treatment.focus == "Ectoparasitoid exclusion" & ectos < 1 | Treatment.focus == "Control") %>% # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
  mutate(Foodweb = ifelse(Treatment.focus=="Control","Complex","Simple"),
         gall_survival = as.numeric(ifelse(pupa > 0, 1, 0)),
         egg_parasitoid = as.numeric(ifelse(platy > 0, 1, 0)),
         sc.Diam = as.numeric(scale(Gall_Height_mm)), 
         sc.log.Clutch = as.numeric(scale(log(gall_individuals))), 
         sc.log1p.Pref = as.numeric(scale(log1p(Density_per_100_shoots)))) 
dim(gall_selection.df)

table(gall_selection.df$Treatment.focus, gall_selection.df$ectos)

## Create separate datasets for quantifying selection gradients ----
#control_df <- gall_selection.df %>%
#  filter(Treatment.focus == "Control") #%>%
  #mutate(sc.gall_size = as.numeric(scale(Gall_Height_mm)), 
  #       sc.clutch_size = as.numeric(scale(gall_individuals)), 
  #       sc.log.clutch_size = as.numeric(scale(log(gall_individuals))), 
  #       sc.female_preference = as.numeric(scale(Density_per_100_shoots)),
  #       sc.log1p.female_preference = as.numeric(scale(log1p(Density_per_100_shoots))))
#dim(control_df)

#treatment_df <- gall_selection.df %>%
#  filter(Treatment.focus == "Ectoparasitoid exclusion", ectos < 1) #%>%  # excluding any larva that were parasitized by an ectoparasitoid in the exclusion experiment.
#dim(treatment_df)
  #mutate(sc.gall_size = as.numeric(scale(Gall_Height_mm)), 
  #       sc.clutch_size = as.numeric(scale(gall_individuals)), 
  #       sc.log.clutch_size = as.numeric(scale(log(gall_individuals))), 
  #       sc.female_preference = as.numeric(scale(Density_per_100_shoots)),
  #       sc.log1p.female_preference = as.numeric(scale(log1p(Density_per_100_shoots))))

#mean(filter(gall_selection.df, Treatment.focus == "Ectoparasitoid exclusion")$ectos) # less than 3% of larva were parasitized by an ectoparasitoid in the exclusion experiment.

@@

## Test for multicollinearity (from: https://jonlefcheck.net/2012/12/28/dealing-with-multicollinearity-using-variance-inflation-factors/)
vif.mer <- function (fit) {
    ## adapted from rms::vif

    v <- vcov(fit)
    nam <- names(fixef(fit))

    ## exclude intercepts
    ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
    if (ns > 0) {
        v <- v[-(1:ns), -(1:ns), drop = FALSE]
        nam <- nam[-(1:ns)]
    }

    d <- diag(v)^0.5
    v <- diag(solve(v/(d %o% d)))
    names(v) <- nam
    v
}

@@

## Calculate selection gradients following method of Janzen and Stern 1998, Evolution
Beta_avg_grad <- function(.) {
  mean_fitness <- mean(predict(., type = "response")) 
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness ## need to divide by mean fitness to put on relative fitness scale.
}

@@

## Compute the matrix describing the curvature of the fitness landscape
curvature_fitness.landscape <- function(gammas, # matrix of nonlinear and correlational selection gradients 
                                        betas){ # vector of directional selection gradients
  gammas - betas %*% t(betas)
}

@@

## Number of bootstrap replicates
n_boots_analysis <- 30 #1000
n_boots_plots <- 100

@@

@@

@@

@@
